LCS Tool: A Computational Platform for Lagrangian Coherent Structures

### 2014-10-07 main.tex
Remove \mathrm from formula
Change writing style in a few sentences

@@

Here we present a simple algorithmic introduction to geodesic LCS detection without details on the the underlying mathematics.
% FIXME JCompSci editor: Replace GitHub link with http://runmycode.org/
% FIXME Reviewer #3, comment #4: Add documentation within the github repository to show some reference to the APIs and how users may start to run a his/her own application.
% FIXME Reviewer #4: Please make sure the code is downloadable once the paper is published
We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is publicly available at: https://github.com/jeixav/LCS-Tool}.
This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows. The examples we present here have been programmed into demonstration scripts.

% FIXME Reviewer #3, comment #3: The comparison with other publicly available LCS detection software should include more details. It contains a very high-level summary of LCS tool being an "all-purpose numerical engine". That part should be revisited and become more concrete.
The list of publicly available software for LCS detection further
includes the following packages:
\begin{itemize}
\item \emph{MANGEN}\citep{lekien03:_time} calculates FTLE and advects material curves in two dimensional velocity fields. It also includes a graphical user interface and uses MPI for parallel calculations.
% FIXME Cannot cite LCS MATLAB Kit without a journal reference
% \item \emph{LCS MATLAB Kit}\citep{dabiri09:_lmk} calculates the FTLE from velocity datasets. Includes a graphical user interface.
\item \emph{Newman}\citep{toit10:_trans} calculates the FTLE in $N$ dimensions. Assists ridge extraction of FTLE fields. Supports analytic  and dataset based velocity definitions.
% FIXME Cannot cite FlowVC without a journal reference
% \item \emph{FlowVC}\citep{shadden10:_flowvc} is a general purpose LCS platform for two and three dimensional datasets. Parallel calculations are supported with Open\-MP, CUDA and OpenCL.
% FIXME FlowTK
% https://github.com/FlowPhysics/FlowTK
% doi:10.1007%2F978-3-319-04099-8_13
\end{itemize}
These packages primarily target the automated generation of FTLE plots to aid the visual assessment of hyperbolic LCSs in the flow. The purpose of LCS Tool and this article is to provide an all-purpose numerical engine for the geodesic extraction of LCSs as parametrized material curves in unsteady two-dimensional flows, and extend the scope of such extraction to elliptic LCSs.

@@

The common first step in calculating elliptic and hyperbolic LCSs is the computation of the Cauchy-Green strain tensor field $C_{t_0}^t(x_0)$, as defined in \cref{eq:CG}. The function performing this calculation in LCS Tool is \lstinline!eig_cgStrain!. The main steps executed by this function are enumerated in \cref{t:Cauchy-Green algorithm}, while \cref{t:eig_cgStrain syntax} summarizes the syntax of \lstinline!eig_cgStrain!.

@@

The velocity is given by $v(x,y,t) = (-\partial_y \psi, \partial_x \psi)$ where
\begin{gather*}
\psi(x,y,t) = \psi_0(x,y) + \psi_1(x,y,t),\\
\psi_0(x,y) = c_3 y - U L_y \tanh\frac{y}{L_y} + \epsilon_3 U L_y \mathrm{sech}^2\frac{y}{L_y} \cos k_3 x,\\
\psi_1(x,y,t) = U L_y \mathrm{sech}^2\frac{y}{L_y} \mathrm{Re}\left[ \sum_{n=1}^2 \epsilon_n f_n(t) e^{i k_n x}\right].
\end{gather*}
As a forcing function, we choose a solution running on the chaotic attractor of the damped and forced Duffing oscillator, specifically
\begin{gather*}
\od{\phi_1}{t} = \phi_2,\\
\od{\phi_1}{t} = -0.1 \phi_2 - \phi_1^3 + 11 \cos(t),\\
f_{1,2}(t) = 2.625 \times 10^{-2} \phi_1(t/6.238 \times 10^5)
\end{gather*}
The parameter values we use are: $U = 62.66$, $c_2 = 0.205 U$, $c_3 = 0.461 U$, $L_y = 1.77 \times 10^6$, $\epsilon_1 = 0.0075$, $\epsilon_2 = 0.04$, $\epsilon_3 = 0.3$, $L_x = 6.371 \times 10^6 \pi$, $k_n = 2 n \pi/L_x$, $\sigma_1 = 0.5 k_2 (c_2 - c_3)$, $\sigma_2 = 2 \sigma_1$.

>>>>

\DeclareMathOperator{\sech}{sech}

@@

Here we present a simple algorithmic introduction to geodesic LCS detection.
% FIXME JCompSci editor: Replace GitHub link with http://runmycode.org/
% FIXME Reviewer #3, comment #4: Add documentation within the github repository to show some reference to the APIs and how users may start to run a his/her own application.
% FIXME Reviewer #4: Please make sure the code is downloadable once the paper is published
We then review the implementation of this approach in a computational engine called LCS Tool\footnote{LCS Tool is available at: http://github.com/jeixav/LCS-Tool}. This engine is a library of MATLAB functions that extract LCSs from two-dimensional unsteady flows. The examples we present have been programmed into demonstration scripts.

% FIXME Reviewer #3, comment #3: The comparison with other publicly available LCS detection software should include more details. It contains a very high-level summary of LCS tool being an "all-purpose numerical engine". That part should be revisited and become more concrete.
Software for LCS detection includes the following packages:
\begin{itemize}
\item \emph{MANGEN}\citep{lekien03:_time} calculates FTLE and advects material curves in two dimensional velocity fields. It also includes a graphical user interface and uses MPI for parallel calculations.
% FIXME Cannot cite LCS MATLAB Kit without a journal reference
% \item \emph{LCS MATLAB Kit}\citep{dabiri09:_lmk} calculates the FTLE from velocity datasets. Includes a graphical user interface.
\item \emph{Newman}\citep{toit10:_trans} calculates the FTLE in $N$ dimensions. Assists ridge extraction of FTLE fields. Supports analytic  and dataset based velocity definitions.
% FIXME Cannot cite FlowVC without a journal reference
% \item \emph{FlowVC}\citep{shadden10:_flowvc} is a general purpose LCS platform for two and three dimensional datasets. Parallel calculations are supported with Open\-MP, CUDA and OpenCL.
% FIXME FlowTK
% https://github.com/FlowPhysics/FlowTK
% doi:10.1007%2F978-3-319-04099-8_13
\end{itemize}
These packages primarily target the automated generation of FTLE plots to aid the visual assessment of hyperbolic LCSs in the flow. The purpose of LCS Tool and this article is to provide an all-purpose numerical engine for the geodesic extraction of LCSs as parametrized material curves in unsteady two-dimensional flows, and extend the scope of such extraction to elliptic LCSs.

@@

The first step in calculating elliptic and hyperbolic LCSs is the computation of the Cauchy-Green strain tensor field $C_{t_0}^t(x_0)$, as defined in \cref{eq:CG}. The function performing this calculation in LCS Tool is \lstinline!eig_cgStrain!. The main steps executed by this function are enumerated in \cref{t:Cauchy-Green algorithm}, while \cref{t:eig_cgStrain syntax} summarizes the syntax of \lstinline!eig_cgStrain!.

@@

The velocity is given by $v(x,y,t) = (-\partial_y \psi, \partial_x \psi)$ where
\begin{gather*}
\psi(x,y,t) = \psi_0(x,y) + \psi_1(x,y,t),\\
\psi_0(x,y) = c_3 y - U L_y \tanh\frac{y}{L_y} + \epsilon_3 U L_y \sech^2\frac{y}{L_y} \cos k_3 x,\\
\psi_1(x,y,t) = U L_y \sech^2\frac{y}{L_y} \Re\left[\sum_{n=1}^2 \epsilon_n f_n(t) e^{i k_n x}\right].
\end{gather*}
As a forcing function, we choose a solution running on the chaotic attractor of the damped and forced Duffing oscillator, specifically
\begin{gather*}
\od{\phi_1}{t} = \phi_2,\\
\od{\phi_1}{t} = -0.1 \phi_2 - \phi_1^3 + 11 \cos(t),\\
f_{1,2}(t) = 2.625 \times 10^{-2} \phi_1(t/6.238 \times 10^5)
\end{gather*}
The parameter values we use are: $U = 62.66$, $c_2 = 0.205 U$, $c_3 = 0.461 U$, $L_y = 1.77 \times 10^6$, $\epsilon_1 = 0.0075$, $\epsilon_2 = 0.04$, $\epsilon_3 = 0.3$, $L_x = 6.371 \times 10^6 \pi$, $k_n = 2 n \pi/L_x$, $\sigma_1 = 0.5 k_2 (c_2 - c_3)$, $\sigma_2 = 2 \sigma_1$.
